---
title: "Basic Plotting and Summary"
author: "Deependra Dhakal"
date: "May 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To make a simple frequency table that summarizes a categorical variable

```{r}
variablename <- c(rep("ok", 9), "not-ok", "ok", "not-ok", "not-ok", rep("ko", 4))
table(variablename)
```

To construct a bar diagram out of frequency table

```{r}
freqtabledemo <- table(variablename)
barplot(freqtabledemo)
barplot(freqtabledemo, # freqtabledemo is better when expressed in percentage
        main="TITLE", xlab="x-axislabel", ylab="y-axixlabel",
        las=0, # to change the style of axis label. possible values are 0, 1, 2 and 3.
        names.arg=c("name1", "name2", "name3"), # a vector of names to be plotted below each bar or group of bars.
        horiz=TRUE) # default is horiz=FALSE
```

## To make a pie chart

```{r}
pie(freqtabledemo)
```

To modify the plotting characters size

```{r}
plot(x, y, main="nameit", cex = 0.5)
```

"cex=0.5" Halves the size of plotting char than what we originally get. Alter "cex.main" for title size change, "cex.lab" for x and y labels, and "cex.axis" for x and y axis, respectively.

To modify the font settings

```{r}
plot(x, y, main="nameit", font.main = 3) # changing font type "3" gives italicized
```

Change the plotting char type by using pch argument

```{r}
plot(x, y, main = "nameit", pch = "w") # plotting character will be shaped "w".
```

To use different characters with reference to a different categorical variable, "point" function is used.

```{r}
plot(x[cat_var == "cat1"], y[cat_var == "cat1"], pch = "cat1", col = 1, main = "Different points")
points(x[cat_var == "cat2"], y[cat_var == "cat2"], pch = "cat2", col =2)
```

To display more than one plot in the screen simultaneously use "mfrow" or "mfcol" argument in par

To modify the y- and x-axis, use:

```{r}
plot(x, y, main = "nameit", axes = F)# will remove the axes
axis(side = 1, at = num_vector, labels = char_vector)#side = 1 corresponds to the x-axis.
box()#to make box reappear
```


## Summary

Let's use the percapita dataset in data subdirectory

```{r}
percapita <- read.csv("./data/percapita.csv", header = T, sep = ",")
```

Column means, median

```{r}
colMeans(percapita[,2:9])
attach(percapita)
mean(Yr_2010, trim = .5)
apply(percapita[2:9], 2, median)
apply(percapita[2:9], 2, quantile) # default for number of quantiles is five
sort(table(Yr_2010))
table(Yr_2010)
```

Using "prettyR" package mode for a vector could be calculated by function

```{r}
# require(prettyR)
# Mode(Yr_2010)
```

Measuring location via standard scores

z-score is calculated by the function scale

```{r}
zYr_2010 <- scale(Yr_2010)
head(zYr_2010)
```

Testing the assumption that z-score is standardized across the variable so that average of z-scores is "0" and sd is "1" could be done by calculating

```{r}
mean(zYr_2010)
apply(zYr_2010, 2, sd)
```

## Measuring variability
Try inbuilt `range()` function of R. Or, Manually write a function for range difference calculation:
For testing, we read in weights dataset from data subdirectory as a vector.

```{r}
weights <- as.vector((read.csv("./data/weights.csv", header = T))$x)
range.difference <- function(x) max(x) - min(x)
range.difference(weights)

```

Median and mean absolute deviations provide an idea of distribution of data with all of them equal or similar, that implies normal distribution.

```{r}
mad(weights)
mad(weights, center = mean(weights))
sd(weights)
hist(weights)
```

Interquartile range is given by
```{r}
IQR(weights)
```

Create "coefficient of variation" function

```{r}
CV <- function(x) sd(x)/mean(x)
CV(weights)
```

Correlation and covariation

```{r}
sales <- read.csv("./data/sales.csv", header = T)
head(sales)
```

An elegant way to rename the dataframe columns(i.e., the "names" attribute)

```{r}
attributes(sales)
class(attributes(sales))
attributes(sales)$names
attributes(sales)$names <- c("Region", "Advertising", "Sales")
```

Unlike variance, the covariance can be "+ve", "-ve" or "0"

```{r}
attach(sales)
cov(Advertising, Sales)
```

Try and make a correlation coefficient function

```{r}
corr.fun <- function(x, y)
(cov(x, y)/(sd(x)* sd(y)))
corr.fun(Advertising, Sales)
```

Use function: cor() for creating matrix of correlation coefficients, using variables from dataframe.

## Measuring Symmetry(or lack thereof)
Calculate skewness from functio skew() and kurtosi() both from the package = "psych".

